---
title: "Data Visualization Project 2"
author: 
- name: Alexander Jaenisch & Rafaela Neff
  affiliation: Beuth University of Applied Science Berlin
date: "Summer Term 2018"
output: 
  html_document: 
    fig_caption: yes
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr) #to set up options and use kable function
opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE) #(echo = TRUE) would include all code
read_chunk('src/animalBites.R')
```

# Introduction

In this excersise we examine animal bites registered in the U.S from

The repository can be found on [Github](https://github.com/amidisturbing/datavis2018).

## The data

**Data source: https://data.louisvilleky.gov/dataset/animal-bites (June 2018)**

```{r part1}
```

The original dataset contains 19826 rows and 18 columns.

The column *head_sent_date* shows the most missing values since not every head of the animal which was registered is cut off and send to the department of health care.

According to the current Data Manager and Performance Partner of the Louisville Metro Department of Public Health and Wellness, Gerald (Jerry) Kaforski, the column *FollowupYNDesc* was added by mistake.

Taking a closer look at the column *bite_date*, we find some the year 1952 until the year 5013.

## Data preparation

The missing values of the dataset were converted to **NA**. 

The values in "bite_date" were limited to the period between the 5th of May 1985 and the 21st of June 2018, what took also care of the impossible values in the dataset.

# Analysis

At first, we want to visualize the first five Species which are involved in Animal Bites in the given time period.

The plot shows that there were more than 15000 dogs, about 35000 a few bats and raccoons and some other animals in the Top 5.

```{r fig.cap="The plot shows that there were more than 15000 dogs, about 35000 a few bats and raccoons and some other animals in the Top 5.", animalBitesPerSpecies}
```

```{r fig.cap="", pieChartForBites}
```

Since dogs are much more likely to be in contact with humans, this comes as no surprise.
For our further Analysis, we will take a closer look at the data we have specifically about dog bites.

Visualizing the dog breeds of the animal bite dataset we clearly see, that pit bulls are on the very top of the ranking.

```{r fig.cap="", dogBitesPerBreed}
```

```{r fig.cap="", genderOfDogsInvolvedInBitesBarplot}
```

```{r fig.cap="", genderOfDogsInvolvedInBites}
```

```{r fig.cap="", breedOfMaleDogsInvolvedInBites}
```

```{r fig.cap="", breedOfFemaleDogsInvolvedInBites}
```

```{r fig.cap="", breedOfUnknownDogsInvolvedInBites}
```

```{r fig.cap="", dogBitesPerMonth}
```

```{r fig.cap="", dogBitesPerDay}
```

```{r fig.cap="", dogBitesPerBreedInTheHead}
```

```{r fig.cap="", dogBitesPerBreedInTheBody}
```

```{r fig.cap="", pitbullBitesPerYear}
```

```{r fig.cap="", catBitesPerGenderAsWafflechart}
```

```{r fig.cap="", catBitesPerYear}
```

```{r fig.cap="", catBitesPerMonth}
```

```{r fig.cap="", catBitesPerDay}
```

```{r fig.cap="", otherSpeciesExcludingDogsAndCats}
```

```{r fig.cap="", otherAnimalBitesPerYear}
```

```{r fig.cap="", otherAnimalBitesPerMonth}
```

```{r fig.cap="", otherAnimalBitesPerDay}
```
