---
title: "Data Visualization Project 2"
author: 
- name: Alexander Jaenisch & Rafaela Neff
  affiliation: Beuth University of Applied Science Berlin
date: "Summer Term 2018"
output: 
  html_document: 
    fig_caption: yes
    highlight: tango
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(knitr) #to set up options and use kable function
opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE) #(echo = TRUE) would include all code
read_chunk('src/animalBites.R')
```

# Introduction

In this excersise we examine animal bites registered in the U.S from

The repository can be found on [Github](https://github.com/amidisturbing/datavis2018).

## The data

**Data source: https://data.louisvilleky.gov/dataset/animal-bites (June 2018)**

First we want to get an idea, how our data looks like and of what quality it is.

```{r part1}
```

The original dataset contains 19826 rows and 18 columns.

The column *head_sent_date* shows the most missing values since not every head of the animal which was registered is cut off and send to the department of health care.

According to the current Data Manager and Performance Partner of the Louisville Metro Department of Public Health and Wellness, Gerald (Jerry) Kaforski, the column *FollowupYNDesc* was added by mistake.

Examining the data by hand, it stood out that the column "bite_date" has some very old entries from the year 1952 were other animal breeds were likely to live in households then it may be the case today.  
It also held 2 impossible entries from the year 5013, which might be a typo meaning  2013.

## Data preparation

The missing values of the dataset were converted to **NA**. 

The values in "bite_date" were limited to the period between the 5th of May 1985 and the 21st of June 2018, what took also care of the impossible values in the dataset.

# Analysis

At first, we want to visualize the first five Species which are involved in Animal Bites in the given time period.

The plot shows that there were more than 15000 dogs, about 35000 a few bats and raccoons and some other animals in the Top 5.

```{r fig.cap="The plot shows that there were more than 15000 dogs, about 35000 a few bats and raccoons and some other animals in the Top 5.", animalBitesPerSpecies}
```

```{r fig.cap="", pieChartForBites}
```

Since dogs are much more likely to be in contact with humans, this comes as no surprise.
For our further Analysis, we will take a closer look at the data we have specifically about dog bites.


##Dog Bites

We are most interessted in the Species we have the most information available. To analyse our visualizations we add the opinion of Steffen Baldwin, expressed in his article *The Lies, Damn Lies and Statistics Behind Dog Bites*. It was published in 2015 at the URL 
https://www.huffingtonpost.com/steffen-baldwin/the-lies-damn-lies-and-st_b_8112394.html.


```{r fig.cap="Visualizing the dog breeds of the animal bite dataset we clearly see, that pit bulls are on the very top of the ranking.", dogBitesPerBreed}
```


```{r fig.cap="", genderOfDogsInvolvedInBitesBarplot}
```

```{r fig.cap="", genderOfDogsInvolvedInBites}
```

```{r fig.cap="", breedOfMaleDogsInvolvedInBites}
```

```{r fig.cap="", breedOfFemaleDogsInvolvedInBites}
```

```{r fig.cap="", breedOfUnknownDogsInvolvedInBites}
```

```{r fig.cap="The most reports in our dataset are from the period  2010 until 2017. Here we encounter around 2000 entries per year.", dogBitesPerYear}
```

The fast growth of the archived reports could be due to the awareness of the importance of data as well as the fast development in the field of IT.

```{r fig.cap="There are more dog bites reported in the warm months.", dogBitesPerMonth}
```

Since it is much more likely that dogs get in contact with strangers during the warm sessions, this adaption seems to be quite reasonable. The chances of getting bitten by a dog are much higher if there are more unknown people around the animal. Also, the more dogs you meet, the higher your chances are to get bitten. 

```{r fig.cap="The day of the month has no influence on the number of reported dog bites.", dogBitesPerDay}
```

The distribution of the gender of the recorded dogs that attacked someone stays equal during the investigated time intervals.

```{r fig.cap="", dogBitesPerBreedInTheHead}
```

```{r fig.cap="", dogBitesPerBreedInTheBody}
```

```{r fig.cap="", pitbullBitesPerDay}
```

By just looking at the charts it might seem obvious to conclude that Pitbulls are most likely to bite human beings, no matter of the gender of the dog, nor if the victim was bitten in head or body.

Now we will take into the account the earlier mentioned arguments of Steffen Baldwin. According to his article the questions 

A, What is a Pit Bull?

and

B, How are pit bulls identified?

are answered as follows:

**What is a Pit Bull?:** It is commonly accepted that “pit bull” is not a breed but a loosely defined and general category. Definition of this category varies depending upon the source. Any blocky headed dog, or any mix of breeds that is between 35 and 100 pounds and upwards of 30 individual dog breeds may currently fall in this broad category through the use of visual breed identification. 

**How are pit bulls identified?:** Visual breed identification, or the process of tilting your head 30 degrees to the right or left and guessing based on a few, basic traits and the inability to automatically identify the dog as something else, is how pit bulls are identified.

Since the most dogs living in American households are mixed Breeds 


## Cat Bites

```{r fig.cap="", catBitesPerGenderAsWafflechart}
```

```{r fig.cap="", catBitesPerYear}
```

```{r fig.cap="", catBitesPerMonth}
```

```{r fig.cap="", catBitesPerDay}
```

```{r fig.cap="", otherSpeciesExcludingDogsAndCats}
```

```{r fig.cap="", otherAnimalBitesPerYear}
```

```{r fig.cap="", otherAnimalBitesPerMonth}
```

```{r fig.cap="", otherAnimalBitesPerDay}
```
